Molecular Gas Dynamics of the lensed wet-merger RXSJ1131-1123 at $z$=0.654

### 2016-05-18 WRITEUP/ms.tex
type-setting, added stuff to lens model sec, not complete

\keywords{ISM: molecular --
          infrared: galaxies --
          galaxies: starburst --
          galaxies: evolution}

@@

% poor phase note
% /Users/admin/Research/RXJ1131/CARMA/imagingD23/checkQuality.csh
%
% calflux source=1127-189 in=$MIRCAT/FluxSource.cat device=/xs
% Flux of: 1127-189  14FEB04.50 at 227.0 GHz:  0.65 Jy; rms: 0.10 Jy
% Extrapolate to 215.673 GHz with spectral index = -0.986 --> 0.684 Jy
% MIRIAD from bootstrap: Median Flux @ 215.673 GHz:     0.710
Given that the phase calibrator used for the first track was faint and was
observed under poor weather conditions and that the phase calibrator used for
the second track was far from our target source, the phase calibration is
subpar, with an rms scatter $\sim$60\degr. We thus conservatively estimate
a calibration accuracy of $\sim$45\% based on the flux scale uncertainties,
gain variation over time, and the phase scatter on the calibrated data. We
therefore treat its line intensity with caution and refrain ourselves from 
using it to derive other physical quantities.

@@

%--------------------------------------------------------------------------
%                                Results
%--------------------------------------------------------------------------
\section{Results}
\subsection{\bco Emission} %DONE
We dynamically resolved the \bco line emission toward the background galaxies
at $\gtrsim$27$\sigma$ significance, confirming the redshift at $z_{\rm CO}$ =
0.65370 $\pm$ 0.0005. A highly asymmetric double-horned line profile is
detected as shown in \Fig{CO21spec}. Fitting a double Gaussian results in peak
flux densities of 75.3$\pm$2.6 and 24.0$\pm$2.0 mJy, and a FWHM of
179$\pm$9 \kms and 255$\pm$28 \kms, respectively. The peaks are separated by
$\Delta v_{\rm sep}$ = 220 \kms. The integrated line flux is 24.1\,$\pm$\,2.3 Jy \kms. % see 15May16/

@@

% moment 0 map: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth, first and second moment maps shown in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube, over $\Delta v$$\sim$750 \kms.
The first and second moment maps are clipped at 3$\sigma$.
The peak flux density is 8.12 $\pm$ 0.30 Jy\,\kms\,beam\pmOne in the intensity-integrated
map. The deconvolved source size is estimated to be 5\farcs1$\pm$0\farcs72 $\times$ 3\farcs72$\pm$0\farcs66 
and thus the emission is resolved over $\sim$2.2 beams. 
While the lensed emission is not strictly distributed as 2D-Gaussian, 
the fit recovers the line intensity enclosed within the emitting
region, we therefore take this as an estimate on the extent of the lensed
emission. On the other hand, if we assume the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, including the partially complete ``Einstein Ring" and
the lensed knots (see \Fig{CO21mom}).

\begin{figure}[!htbp]
\centering
\includegraphics[trim=0 5 0 15, clip, width=0.232\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=0 5 0 0, clip, width=0.22\textwidth]{../Figures/F555W_REDBLUE.eps}
\caption{
An overlay of the velocity-integrated \bco emission on the HST V$-$band (F555W)
image. The contours start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne. The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295. {\it Right:} The contours
are color-coded to represent the red and blue wings of the emission.
\label{fig:CO21mom}}
\end{figure}

@@

\begin{figure*}[tbph]
\centering
\includegraphics[width=1.0\textwidth]{../Figures/co_channel_maps.eps}
\caption{
Channel maps of \pdbi \bco toward RXJ1131 in 21.5\,\kms resolution. The
Black crosses indicate the position of the lensed knots (AGN emission,
which are components ABCD in C06). The central white-filled
star indicates the position of the foreground lensing galaxy (component G
in C06), as detected in the HST image. Contours start and increment at steps of
$\pm$3$\sigma$. The beam is denoted in the bottom right panel. \label{fig:chanmap}}
\end{figure*}

@@

\subsection{Origin of the \bco emission} \label{sec:origin} %DONE
Evidently, the CO emission is co-spatial with the lensed optical emission in
the background AGN host galaxy. The question is then whether the optically
faint companion also contributes to the CO flux and thus contains a
non-negligible amount of molecular gas.
Indeed, based on our lens model presented in \Sec{lensmodel}, we find it
highly likely that part of the CO emission originates from
the companion. We thus classify RXJ1131 as a ``wet-wet'' merger.
However, we cannot quantify this with a mass ratio given the
spatial resolution. This will be verified in future higher-resolution observations. In the subsequent sections, we will interpret RXJ1131
using this scenario (i.e. a ``wet-wet'' merger).

\subsection{\bco Kinematics} % DONE
A clear velocity gradient and an unusually high
velocity dispersion ($\gtrsim$400\kms) near the central region
is seen in \Fig{CO21highO}. While beam smearing is inevitably the
dominant factor in the observed velocity dispersion
at the spatial resolution of this data, the exceedingly
high velocity dispersion hints
at potential perturbations from the AGN, or internal turbulence due to
interactions with the companion, and/or instability due to the large gas
content. A velocity gradient across the source plane in \Fig{model} is also
found in our lens model. This further supports the idea of a kinematically-
ordered galaxy, but its emission has been lensed differentially.
In this scenario, RXJ1131 is a disrupted disk galaxy hosting an optically
bright quasar and is in the process of merging.
We will defer the discussion of the implications to \Sec{dynamics},
where we analyze the dynamics of the system in the source plane.

@@

\begin{figure}[!Htbp]
\centering
\includegraphics[width=0.455\textwidth]{../Figures/coOverlay.eps}
\caption{\carma \cco line profile (solid) without continuum subtraction is
over-plotted on the continuum-subtracted \bco line profile (dashed).
The velocity scale is with respect to $z$=0.6537, which corresponds to the
dynamical center of the \bco line. The spectral resolution for \cco and \bco
is 35.8 \kms and 21.5 \kms, respectively.
 \label{fig:co32spec}}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.23\textwidth]{../Figures/F555W_ContPdBI.eps}
\hspace{-0.5em}
\includegraphics[width=0.235\textwidth]{../Figures/F555W_ContVLA.eps}
\caption{{\it Left:} An overlay of the 2\,mm continuum emission on the optical image. 
{\it Right:} VLA 5\,GHz continuum emission is overlaid on the optical image. 
Contours in both images start and increment at steps of 
$\pm$3$\sigma$, where $\sigma_{\rm 2mm}$\,=\,0.082 mJy beam\pmOne and 
$\sigma_{\rm 5GHz}$\,=\,13 $\mu$Jy beam\pmOne in the left and right panel, respectively.
The central crosses indicate the centroid of the foreground galaxy,
as detected in the optical image. The synthesis beams are shown in the bottom left corner of each panel.
\label{fig:cont}}\vspace{0.51em}
\end{figure}

@@

%--------------------------------------------------------------------------
%                                Analysis
%--------------------------------------------------------------------------
\section{Analysis}
\subsection{Lens modeling} \label{sec:lensmodel}
The spectrally resolved lensed emission allows us to probe dynamical structures on smaller spatial scales than otherwise possible.

With the high SNR and spectral resolution, we model different kinematic components using our lensing code \uvmcmcfit.

% describe model setup.

@@

\begin{figure*}[tbph]
\centering
\includegraphics[width=0.85\textwidth]{../Figures/PostageStampModel.eps}
\includegraphics[width=0.85\textwidth]{../Figures/PostageStampResiduals.eps}
\caption{
Lens model of channel widths $\sim$ 100\kms.
Make this bigger in the future
Channel maps of the CO emission (red) overlaid on our best-fit lens models (grayscale). The foreground lensing galaxy is represented by a black dot. The reconstructed source morphology (magenta ellipses) is also suggestive of a disk.
\label{fig:model}}
\end{figure*}

@@

\subsection{SED modeling} \label{sec:SED}

We extrapolated a power-law to 24 \micron
to decompose the flux at 24 \micron in MIPS.

\subsection{ISM properties}
\subsection{Gas Mass}
\subsection{Source Extent}
Compared to HST:

>>>>

\keywords{ISM: molecular --
          infrared: galaxies --
          galaxies: mergers --
          galaxies: starburst --
          galaxies: evolution}

@@

% poor phase note
% /Users/admin/Research/RXJ1131/CARMA/imagingD23/checkQuality.csh
%
% calflux source=1127-189 in=$MIRCAT/FluxSource.cat device=/xs
% Flux of: 1127-189  14FEB04.50 at 227.0 GHz:  0.65 Jy; rms: 0.10 Jy
% Extrapolate to 215.673 GHz with spectral index = -0.986 --> 0.684 Jy
% MIRIAD from bootstrap: Median Flux @ 215.673 GHz:     0.710
Given that the phase calibrator used for the first track was faint and was
observed under poor weather conditions and that the phase calibrator used for
the second track was far from our target source, the phase calibration is
subpar, with an rms scatter $\sim$60\degr. We thus conservatively estimate
a calibration accuracy of $\sim$45\% based on the flux scale uncertainties,
the gain variations over time, and the phase scatter on the calibrated data. We
therefore treat its line intensity with caution and ensure that any physical interpretation
of this system (and thus implications and conclusions of this paper) do not rely on this quantity.

@@

%--------------------------------------------------------------------------
%                                Results
%--------------------------------------------------------------------------
\section{Results}
\subsection{\bco Emission} \label{sec:CO21} %DONE
We dynamically resolved the \bco line emission toward the background galaxies
at $\gtrsim$27$\sigma$ significance, confirming the redshift at $z_{\rm CO}$ =
0.65370 $\pm$ 0.0005. A highly asymmetric double-horned line profile is
detected as shown in \Fig{CO21spec}. Fitting a double Gaussian results in peak
flux densities of 75.3$\pm$2.6 and 24.0$\pm$2.0 mJy, and a FWHM of
179$\pm$9 \kms and 255$\pm$28 \kms, respectively. The peaks are separated by
$\Delta v_{\rm sep}$ = 220 \kms. The integrated line flux is 24.1\,$\pm$\,2.3 Jy \kms. % see 15May16/

@@

% moment 0 map: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth, first and second moment maps shown in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube, over $\Delta v$$\sim$750 \kms.
The first and second moment maps are clipped at 3$\sigma$.
The peak flux density is 8.12 $\pm$ 0.30 Jy\,\kms\,beam\pmOne in the intensity-integrated
map. The deconvolved source size is estimated to be 5\farcs1$\pm$0\farcs72 $\times$ 3\farcs72$\pm$0\farcs66
and thus the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as 2D-Gaussian,
the fit recovers the line intensity enclosed within the emitting
region, we therefore take this as an estimate on the extent of the lensed
emission. On the other hand, if we assume the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein Ring" and
the lensed knots (see \Fig{CO21mom}).

\begin{figure}[!htbp]
\centering
\includegraphics[trim=0 5 0 15, clip, width=0.232\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=0 5 0 0, clip, width=0.22\textwidth]{../Figures/F555W_REDBLUE.eps}
\caption{
An overlay of the velocity-integrated \bco emission on the HST V$-$band (F555W)
image. The contours start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne. The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295. Right: contours
are color-coded to represent the red and blue wings of the emission.
\label{fig:CO21mom}}
\end{figure}

@@

\begin{figure*}[tbph]
\centering
\includegraphics[width=1.0\textwidth]{../Figures/co_channel_maps.eps}
\caption{
Channel maps of PdBI \bco toward RXJ1131 in 21.5\,\kms resolution.
Black crosses indicate the position of the lensed knots (AGN emission,
which correspond to components ABCD in C06). The central white-filled
star indicates the position of the foreground lensing galaxy (component G
in C06). Contours start and increment at steps of
$\pm$3$\sigma$. The beam is denoted in the bottom right panel. \label{fig:chanmap}}
\end{figure*}

@@

\subsection{Origin of the \bco emission} \label{sec:origin} %DONE
Evidently, the CO emission is co-spatial with the lensed optical emission in
the background AGN host galaxy. The question is then whether the optically
faint companion also contributes to the CO flux and thus contains a
non-negligible amount of molecular gas.
Indeed, based on our lens model presented in \Sec{lensmodel}, we find it
highly likely that part of the CO emission originates from
the companion. We thus classify RXJ1131 as a ``wet-wet'' merger.
However, we cannot quantify this with a mass ratio given the
spatial resolution. This will be verified in future higher-resolution observations.
In the subsequent sections, we will interpret RXJ1131
assuming this classification (i.e. a ``wet-wet'' merger).

\subsection{\bco Kinematics} % DONE
A clear velocity gradient and an unusually high
velocity dispersion ($\gtrsim$400\kms) near the central region
is seen in \Fig{CO21highO}. While beam smearing is inevitably the
dominant factor in the observed velocity dispersion
at the spatial resolution of this data, the exceedingly
high velocity dispersion hints
at potential perturbations from the AGN, or internal turbulence due to
interactions with the companion, and/or instability due to the large gas
content. A velocity gradient across the source plane in \Fig{model} is also
found in our lens model. This further supports the idea of a
kinematically-ordered galaxy, but its emission has been lensed differentially.
In this scenario, RXJ1131 is a disrupted disk galaxy hosting an optically
bright quasar and is in the process of merging.
We will defer the discussion of its implications to \Sec{dynamics},
where we analyze the dynamics of the system in the source plane.

@@

\begin{figure}[!Htbp]
\centering
\includegraphics[width=0.455\textwidth]{../Figures/coOverlay.eps}
\caption{CARMA \cco line profile (solid) without continuum subtraction is
over-plotted on the continuum-subtracted \bco line profile (dashed).
The velocity scale is with respect to $z$=0.6537, which corresponds to the
dynamical center of the \bco line. The spectral resolution for \cco and \bco
is 35.8 \kms and 21.5 \kms, respectively.
 \label{fig:co32spec}}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.23\textwidth]{../Figures/F555W_ContPdBI.eps}
\hspace{-0.5em}
\includegraphics[width=0.235\textwidth]{../Figures/F555W_ContVLA.eps}
\caption{Left: an overlay of the 2\,mm continuum emission on the optical image.
Right: VLA 5\,GHz continuum emission is overlaid on the optical image.
Contours in both images start and increment at steps of
$\pm$3$\sigma$, where $\sigma_{\rm 2mm}$\,=\,0.082 mJy beam\pmOne and
$\sigma_{\rm 5GHz}$\,=\,13 $\mu$Jy beam\pmOne in the left and right panel, respectively.
The central crosses indicate the centroid of the foreground galaxy,
as detected in the optical image. The synthesis beams are shown in the bottom left corner of each panel.
\label{fig:cont}}\vspace{0.51em}
\end{figure}

@@

%--------------------------------------------------------------------------
%                                Analysis
%--------------------------------------------------------------------------
\section{Analysis}
\subsection{Lens modeling} \label{sec:lensmodel}
At the angular resolution of the \bco data, the images are resolved
but not within the multiply lensed emission (i.e., sub-optimal for
performing lens modeling). Nevertheless, the high spectral
resolution of this data provides dynamical information on
spatial scales smaller than the beam (see \Fig{CO21highO}).
Hence, with the high SNR, we reconstruct the intrinsic gas
dynamics by carrying out parametric lens modeling over different
channel slices of the interferometric data using our lensing code
\uvmcmcfit (\citealt{uvmcmcfit15a};see \citealt{Bussmann15a} for details of
the code). Model of each slice in \Fig{model} thus provides
properties on the intrinsic kinematics. The slices
are obtained by binning the data (over the full linewidth
$\sim$700\kms; see \Sec{CO21}) by five channels (i.e., into seven slices)
 to increase the SNR.

The lens mass distribution is modeled using a singular isothermal
ellipsoid (SIE) profile, which is described by five free parameters: the
positional offset in R.A. and Dec. relative to an arbitrary chosen
fixed coordinate in the image, the Einstein Radius, the axial ratio, and the
position angle. We use the VLA radio continuum emission toward
the foreground galaxy to initialize the positional offset. We impose a
uniform prior $\pm$0\farcs05 in both $\Delta$R.A. and $\Delta$Dec., which is
larger than the astrometry uncertainties in the VLA image and is
constrained by the uncertainties provided from previous SIE lens model (C06).
We initialize the Einstein Radius based on the optical SIE lens
model by C06 and impose a uniform prior using their reported $\pm$3$\sigma$.
The sources are modeled using elliptical Gaussian profiles, which are
parameterized by six free parameters: the positional offset in R.A.
and Dec. relative to the lens, the intrinsic flux density, the effective
radius, the axial ratio, and the position angle. The position of each source
is allowed to vary between $\pm$1\farcs5 (i.e., within the Einstein Radius)
and the effective radius is allowed to vary from 0\farcs01$-$2".

% Check with Shane
Our code uses an Markov Chain Monte Carlo (MCMC) approach to sample the
posterior probability distribution function (PDF). In each model, we employ
$\sim$50,000 samples as the initial ``burn-in'' stage to stabilize the Markov
chains (which we then discard). We use the final $\sim$5,000
samplings to determine the posterior probability distributions. Here, we
identify the best-fit model and the quoted uncertainties using the
median and the 68\% confidence intervals in the marginal PDFs.

We first obtain a preliminary lens model for each channel slice independently,
where their lens parameters are allowed to vary and are initialized according
to the aforementioned way. We obtain the final model
by repeating the modeling over each slice but fixing their lens parameters
according to the overall median in the preliminary models
(see \Tab{lens}). This ensures the
lens model of each kinematic component share the same lens profile.
The magnification factor is determined by taking the ratio
between the image plane flux of the model and the source plane flux.
The source positions and magnification factors are listed in \Tab{model}.

% disk and caveat
The location of the reconstructed sources in \Fig{model} are consistent with
of a disk. This in essence demonstrate a velocity field in the source plane.
 While solely based on the lens model would not be sufficient to
draw such conclusion/assertion. Yet, with the observed velocity gradient.
We emphasis that the spatial resolution of is a few arcsec, which means that
although we have high SNR and spectral resolution, this data is not sufficient
to constrain the intrinsic sizes of the lensed galaxies, and thus underpredict
the magnification factors. We will discuss this further in the following
section.

@@

\begin{figure*}[tbph]
\centering
\includegraphics[width=0.85\textwidth]{../Figures/PostageStampModel.eps}
\includegraphics[width=0.85\textwidth]{../Figures/PostageStampResiduals.eps}
\caption{
Lens models of RXJ1131 using \uvmcmcfit on the PdBI \bco data.
Lens model of channel widths $\sim$ 100\kms.
Make this bigger in the future
Top: Channel maps of the PdBI \bco emission (red) overlaid on our best-fit lens models (grayscale). The location of the foreground lensing galaxy is indicated by a black dot and its critical curve is traced by the orange solid line. The locations and morphologies (half-light radii) of the reconstructed sources are
represented by magenta ellipses (see text in \Sec{lensmodel} for caveat). The caustic curves are shown as cyan lines.
The beam of the observations is shown in
the bottom right corner of each panel.
Bottom: Residual images of the best-fit model, obtained by
taking the Fourier transform after subtracting the best-fit model from the
data in the visibility plane.
Contours start at $\pm$3$\sigma$ and increment at steps of
3$\times$2$^n\sigma$, where $n$ is a positive integer.
\label{fig:model}}
\end{figure*}

@@

\subsection{SED modeling} \label{sec:SED}
We extrapolated a power-law to 24 \micron to decompose the flux
at 24 \micron in MIPS.

We fit modified blackbody functions to
to constrain the dust temperature, \fir (rest-frame 42.5-122.5 \micron) luminosity, dust mass (For the mass absorption coefficient we adopt κν = 2.64 m2 kg−1 at 125 μm (rest frame; Dunne et al. 2003).)
We added blah\% calibration uncertainties on the PdBI data.

We check convergence of our chains by requiring that the autocorrelation length for each model parameter is several times less than the number of steps taken during an initial burn-in stage (which are then discarded).
Compared with the version of this code used by Riechers et al. (2013), several improvements have been made. First, we have added the ability to impose upper and lower limits as well as Gaussian priors on the model parameters. These additions are useful for analyzing poorly constrained data, such as when we consider fits to only the SPIRE photometry in Section 3. Second, instead of treating the instrument response as a delta function in wavelength, we now fully model the wavelength response of each instrument. We have applied this improved code to FLS 3, and find that the resulting changes in the model parameters are well within the quoted uncertainties.

% upper limit using decomposed 24 um
% lower limit using fit without 24 um, but do I still want to fit a power-law for this?
The best-fit SEDs are shown in \Fig{SED} and the resulting parameters (inferred from the BLAH of the posterior PDF) are listed in \Tab{SED}.
% parameter estimation problem
For these parameter estimation problems using MCMC to sample,
the full answer is the full joint posterior PDF.
We here report the summary statistics (global mode (where min. chi),
marginal mode (in log file),
posterior median (instead of posterior means; need to compute) and the 1$\sigma$ confidence intervals in the marginal PDF) to reflect the PDF shape.

\subsection{ISM properties}
We correct for lensing magnification using the median among our models ($\mu_{\rm L}$=BLAH).
Assuming a Salpeter initial mass function, we find the SFR of blah using the \fir luminosity after corrected for lensing .
\subsection{Gas Mass}
\subsection{Source Extent}
Compared to HST:
